# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
temp = tempfile()
download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)

data = read.csv(unz(temp, "activity.csv"))
x= is.na(data[1])
data1 = data[!x,]
```
## What is mean total number of steps taken per day?
```{r}
sums = aggregate(steps ~ date , data = data1, FUN = sum)
means = aggregate(steps ~ date , data = data1, FUN = mean)
medians = aggregate(steps ~ date , data = data1, FUN = median)

tot = cbind(means,medians[2])
colnames(tot) = c("date", "mean steps", "median steps")

hist(sums$steps,xlab = "steps", main = "total steps per day") 

```

The Calculated mean and median total number of steps taken per day are.

```{r showtable,results='asis'}
library(xtable)
xt = xtable(tot)
print(xt, type = "html")


```


## What is the average daily activity pattern?
```{r}

avg = aggregate(steps ~ interval , data = data1, FUN = mean)

plot(avg$interval, avg$steps,xlab = "5 sec interval",ylab = "steps",type = "l")
maxx = max(avg[2])
intv = avg[2] == maxx
max1 = avg[intv,]

```
The interval,that contains the maximum number of steps is `r max1[1]`

## Imputing missing values
```{r}
xx = sum(x)

data3 = data[x,]
data4 = data3[2:3]

merged.data <- merge(data3, avg, by="interval")

final = merged.data[c(1,3,4)]
colnames(final) = c("interval","date","steps")
 fin.merg = rbind(final,data1)

sums = aggregate(steps ~ date , data = fin.merg, FUN = sum)
means = aggregate(steps ~ date , data = fin.merg, FUN = mean)
medians = aggregate(steps ~ date , data = fin.merg, FUN = median)

tot = cbind(means,medians[2])
colnames(tot) = c("date", "mean steps", "median steps")

hist(sums$steps,xlab = "steps", main = "total steps per day with na values replaced") 


```

The total number of missing values in the dataset is `r xx`




The new Calculated mean and median of total number of steps taken per day are.

```{r showtabl,results='asis'}
xtt = xtable(tot)
print(xtt, type = "html")


```

Overall the structure of the graph looks unchanged from the orignal graph. The mean and median valuses on each day looks also unchancged, the only thing I notice was that now some new dates are showing up.


## Are there differences in activity patterns between weekdays and weekends?


``` {r}
fin.merg$date <- as.Date(fin.merg$date)
fin.merg=cbind(fin.merg, weekend=0)

for (i in 1:nrow(fin.merg)) {
  if(weekdays(fin.merg[i,"date"])== "Saturday" || weekdays(fin.merg[i,"date"])== "Sunday")
    {
        
    fin.merg[i,"weekend"] = 1
    }
}
weekday = fin.merg[4] == 1
weekday1 = fin.merg[!weekday,]
weekend= fin.merg[weekday,]

par(mfrow = c(2, 1),mar = c(2, 2, 2, 2))

plot(weekday1$interval, weekday1$steps,xlab = "5 sec interval",ylab = "steps",main= "weekday",type='l' )
plot(weekend$interval, weekend$steps,xlab = "5 sec interval",ylab = "steps",main= "weekend",type='l' )




```